Language Bias in a Multilingual Tweet Corpus

### 2018-06-27 paper.tex
More references
Supplement files are links to Zenodo

\newcommand{\sfile}[1]{\texttt{#1}}

@@

%%%%%%%%%
% Problem
%%%%%%%%%
%There are several challenges of building a tweet collection that adequately represents language use in the country.
%
Building a monolingual Latvian collection could be done by harvesting tweets that contain indicative Latvian words, which are not present in other languages, similarly to how it is done for Dutch \cite{sang2013}. However, such an approach is not suitable for tweets in Russian and English, as these languages are widely used outside of Latvia.
%
For the same reason, a TREC-like collection building approach \cite{lin2017rts} of filtering the publicly available stream of tweets by language would not work.
%
A tweet collection based on a curated list of users \cite{SANVICENTE16.465,L14-1642} is effective, but extra care must be taken as a thematic bias might be introduced.
%
An alternative approach would be to retrieve only geo-located tweets \cite{milajevs:2017:BUCC}. Such a collection would neither be biased linguistically because it is not based on a list of keywords, nor it would be biased thematically because it is not based on a list of users. The downside is that a large number of tweets are not geo-located, which makes retrieval incomplete.

@@

\footnotetext{In this paper, by ``following users'', we mean that their user IDs where passed to the Twitter API to collect tweets that they created, retweets of those tweets, or the tweets that are replies by other users to their tweets. Refer to \href{https://developer.twitter.com/en/docs/tweets/filter-realtime/guides/basic-stream-parameters\#follow}{the Twitter documentation} for a complete description. Keep in mind that this is different from the case when a user follows another user.}

@@

To comply with Twitter's terms of service, on \hl{8 May 2018}, the raw tweet data was redownloaded to get rid of deleted tweets. The tweets that originated a retweet were added to the collection. Also, we have noticed a large number of tweets that came from Sweden (probably because of the imprecise \texttt{locations} parameter value), so the tweets which location country code was \texttt{SE} were omitted. This resulted in \hl{1\,242\,943} tweets that formed the final collection presented here.

@@

To investigate the collection further, \hl{16} topics of interests were defined. The collection was manually searched using keywords to get relevant tweets. The topics are:
\begin{itemize}
\item \texttt{LV016:} \textbf{Latvia 100} \ The Centennial Celebration of the Republic of Latvia.
\item \texttt{LV005:} \textbf{Ice hockey} \ Ice hockey.
\item \texttt{LV003:} \textbf{Refugees} \ Refugee crisis in Europe, attitude to immigration and immigrants.
\item \texttt{LV009:} \textbf{Brexit} \ Withdrawal of the United Kingdom from the European Union known as Brexit.
\item \texttt{LV002:} \textbf{The Chronicles of Melanie} \ Mel\=anijas Hronika (The Chronicles of Melanie) is a Latvian movie that was selected for the Foreign-Language Category for Oscars.
\item \texttt{LV004:} \textbf{Blizzard of Souls} \ A Latvian movie Dv\=ese\c{l}u Putenis (Blizzard of Souls) that is in production by the time of writing.
\item \texttt{LV001:} \textbf{Winter} \ Winter, snow and cold weather.
\item \texttt{LV015:} \textbf{May 4} \ The Restoration of Independence Day.
\item \texttt{LV011:} \textbf{May 8-9} \ Remembrance of the end of Word War II (May 8). It is also honored as the Victory Day on May 9. The Europe Day is observed on May 9. Tweets about any of the three events are relevant.
\item \texttt{LV010:} \textbf{Midsummer} \ Midsummer celebration on June 23-24.
\item \texttt{LV014:} \textbf{November 11} \ A memorial day for soldiers who fought for the independence of Latvia.
\item \texttt{LV013:} \textbf{November 18} \ Proclamation Day of the Republic of Latvia.
\item \texttt{LV006:} \textbf{Christmas} \ Christmas.
\item \texttt{LV007:} \textbf{New Year} \ New Year.
\item \texttt{LV008:} \textbf{Olympics} \ The Olympic games and the Latvian team.
\item \texttt{LV012:} \textbf{March 16} \ Remembrance Day of the Latvian legionnaires.\footnotemark{}
\end{itemize}

>>>>

\newcommand{\zenodoBase}{https://zenodo.org/record/1299195}
\newcommand{\sfile}[1]{\href{\zenodoBase/files/#1}{#1}}

@@

%%%%%%%%%
% Problem
%%%%%%%%%
%There are several challenges of building a tweet collection that adequately represents language use in the country.
%
Building a monolingual Latvian collection could be done by harvesting tweets that contain indicative Latvian words, which are not present in other languages, similarly to how it is done for Dutch \cite{sang2013}. However, such an approach is not suitable for tweets in Russian and English, as these languages are widely used outside of Latvia.
%
For the same reason, a TREC-like collection building approach \cite{lin2017rts} of filtering the publicly available stream of tweets by language would not work.
%
A tweet collection based on a curated list of users \cite{SANVICENTE16.465,L14-1642} is effective, but extra care must be taken as a thematic bias might be introduced.
%
An alternative approach would be to retrieve only geo-located tweets \cite{coats_steven_2017,milajevs:2017:BUCC}. Such a collection would neither be biased linguistically because it is not based on a list of keywords, nor it would be biased thematically because it is not based on a list of users. The downside is that a large number of tweets are not geo-located, which makes retrieval incomplete.

@@

\footnotetext{In this paper, by ``following users'', we mean that their user IDs where passed to the Twitter API to collect tweets that they created, retweets of those tweets, or the tweets that are replies by other users to their tweets. Refer to the \href{https://developer.twitter.com/en/docs/tweets/filter-realtime/guides/basic-stream-parameters\#follow}{Twitter documentation} for a complete description. Keep in mind that this is different from the case when a user follows another user.}

@@

To comply with Twitter's terms of service, on \hl{8 May 2018}, the raw tweet data was redownloaded to get rid of deleted tweets. The tweets that originated a retweet were added to the collection. Also, we have noticed a large number of tweets that came from Sweden (probably because of the imprecise \texttt{locations} parameter value), so the tweets which location country code was \texttt{SE} were omitted. This resulted in \hl{1\,242\,943} tweets that formed the final collection presented here.\footnotemark{}

\footnotetext{The supplement files are available at Zenodo: \url{\zenodoBase}. Collected tweet IDs are available as \sfile{collected\_tweets.csv} and the final collection is available as \sfile{rehydrated\_tweets.csv}. The supplement files are licensed under a \href{https://creativecommons.org/publicdomain/zero/1.0/legalcode}{Creative Commons CCZero 1.0 License/Waiver}.}

@@

To investigate the collection further, \hl{16} topics of interests were defined. The collection was manually searched using keywords to get relevant tweets. The topics are:\footnotemark{}
\footnotetext{The topic file is available as \sfile{topics.json.txt}.}
\begin{itemize}
\item \texttt{LV016:} \textbf{Latvia 100} \ The Centennial Celebration of the Republic of Latvia.
\item \texttt{LV005:} \textbf{Ice hockey} \ Ice hockey.
\item \texttt{LV003:} \textbf{Refugees} \ Refugee crisis in Europe, attitude to immigration and immigrants.
\item \texttt{LV009:} \textbf{Brexit} \ Withdrawal of the United Kingdom from the European Union known as Brexit.
\item \texttt{LV002:} \textbf{The Chronicles of Melanie} \ Mel\=anijas Hronika (The Chronicles of Melanie) is a Latvian movie that was selected for the Foreign-Language Category for Oscars.
\item \texttt{LV004:} \textbf{Blizzard of Souls} \ A Latvian movie Dv\=ese\c{l}u Putenis (Blizzard of Souls) that is in production by the time of writing.
\item \texttt{LV001:} \textbf{Winter} \ Winter, snow and cold weather.
\item \texttt{LV015:} \textbf{May 4} \ The Restoration of Independence Day.
\item \texttt{LV011:} \textbf{May 8-9} \ Remembrance of the end of Word War II (May 8). It is also honored as the Victory Day on May 9. The Europe Day is observed on May 9. Tweets about any of the three events are relevant.
\item \texttt{LV010:} \textbf{Midsummer} \ Midsummer celebration on June 23-24.
\item \texttt{LV014:} \textbf{November 11} \ A memorial day for soldiers who fought for the independence of Latvia.
\item \texttt{LV013:} \textbf{November 18} \ Proclamation Day of the Republic of Latvia.
\item \texttt{LV006:} \textbf{Christmas} \ Christmas.
\item \texttt{LV007:} \textbf{New Year} \ New Year.
\item \texttt{LV008:} \textbf{Olympics} \ The Olympic games and the Latvian team.
\item \texttt{LV012:} \textbf{March 16} \ Remembrance Day of the Latvian legionnaires.\footnotemark{}
\end{itemize}
